---
layout: post
title: Analyzing "Brand Ticker" data.
author: flovv
published: false
status: processed
draft: false
tags: R 
---



```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(stringr)
library(lubridate)
library(ggplot2)
library(plyr)
library(grid)
library(gridExtra)
library(stargazer)
require(ggthemes)
require(reshape2)
options(stringsAsFactors = FALSE)
VW <- read.csv("data/BrandData_vw.csv")
stockData <- read.csv("data/StockData_vw.csv")
VW$Date <- as.Date(VW$Date)
stockData$Date <- as.Date(stockData$Date)
```

[The Brand Ticker](http://www.thebrandticker.com) provides brand specific marketing data. On a day to day basis they list the brand value and top 3 associations gathered from news mentions and social data. They provide no documentation on how they process the data or which sources they use. 

Here is short look at their service.
I pulled half a year's data in irregular intervalls from their API using a small r-script. See this [GIST](https://gist.github.com/flovv/30a74925958908372e98)  

Due to some recent attention we start the analysis by looking at the Volkswagen (VW) data.

```{r, echo=FALSE}
tail(VW[,2:(ncol(VW)-1)])
```

The first 6 columns are the brand associations, gathered from news data, or a survey (who knows...?). It's helpful that the flag attributes already labels the associations as positiv/negativ (red or green). The value column is the so-called "Brand Value".


```{r, echo=FALSE, message=FALSE}
ggplot(VW, aes(Date, value))+ geom_line() +theme_economist() + ylab("Brand value in Mio Euro") + xlab("2015")

```

In order to see if the brand value is related to associations, I color-code the brand value with the top (attr_1) association.


```{r, echo=FALSE, message=FALSE}

ggplot(VW, aes(Date, value,group=attr_1, color=attr_1))+ geom_point( size=3) +theme_economist() + labs(color="Top Associations") + ylab("Brand value in Mio Euro") + xlab("2015")

```

For the diesel-gate event, we see that the top association is a rather lagging than predicting variable. The brand value massively drops, the association remains "secure". The association change from "secure" to "polluting" happens  with a distinct lag.
In order to visualize the second and third association, we can use a simple histogram of mentions in the data set.

```{r, echo=FALSE, message=FALSE,  out.width = '900px', dpi=200}
sub <- VW[, c("Date", "attr_1", "attr_2", "attr_3")]
colnames(sub) <- c("Date", "first", "second", "third")
asso <- melt(sub, id.var="Date")


ggplot(asso, aes(value, fill=variable))+ geom_histogram( )  +theme_economist(11) + xlab("Associations") +scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+ labs(fill="Associations") +theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

As before, we can add all associations to the brand value time series.

```{r, echo=FALSE, message=FALSE,  out.width = '900px', dpi=200}
mm <- merge(asso, VW, by="Date", all.x=T)
ggplot(mm, aes(Date, value.y, color=value.x))+ geom_point( size=3) + facet_grid(~variable)+theme_economist(6) + labs(color="Top Associations") + ylab("Brand value in Mio Euro") + xlab("2015")
```

In order to unterstand the brand value variable it bit better, I scraped VW's stock market data from Yahoo. See this [GIST](https://gist.github.com/flovv/50ed4495911576040e34)   

```{r, echo=FALSE, message=FALSE}
m <- merge(VW, stockData, by="Date")

m <- m[, c("Date", "value", "Close")]
colnames(m) <- c("Date", "Brand value in Mio Euro", "Stock price in Euro")
mm <- melt(m, id.var="Date")

ggplot(mm, aes(Date, value))+ geom_line() +theme_economist() + ylab("") + xlab("2015") + facet_wrap(~variable, scale="free_y")
```

Apparently brand value is just a linear function of stock market data and can be regarded as  more or less identical. 

```{r,results='asis', echo=F}
m <- merge(VW, stockData, by="Date")

fit1 <- lm(value~Close, data=m)
fit2 <- lm(value~Close-1, data=m)

stargazer(fit1, fit2, type = "html" ,title = "Brandtickers Brand value has a direct relation to VW's stock price")

```


I find it quite misleading to speak of "Brand Value" and taking stock market data as a reference. Plenty of factors (economic condictions, production costs, ...) have an impact on stock prices, while the value of a brand is related to subjective perceptions. 

